---
title: "Getting started with R and RStudio"
teaching: 0
exercises: 0
---

:::::::::::::::::::::::::::::::::::::: questions 

- How do I navigate the RStudio user interface?
- What do the different panes in RStudio do?
- How do I run commands in the console?
- What is an R script and how do I create one?
- What are packages?
- How do I get help?

::::::::::::::::::::::::::::::::::::::::::::::::


::::::::::::::::::::::::::::::::::::: objectives

After completing this episode, learners should be able to:
- Navigate the user interface of RStudio
- Run commands in the console
- Run commands in an R script
- Use install.packages() to download and install a library package
- Use library() to load the downloaded package
- Use the help(), help.search(), ?, and ?? commands to look up documentation 

::::::::::::::::::::::::::::::::::::::::::::::::


:::::::::::::::::::::::::::::: challenge

### Exercise: Install and load packages

Complete the following code to first install the ggplot2 package and then load the package in your R environment.

Step 1 - Install package

```{r}
install.______('_____')
```

Step 2 - Load package

```{r}
library(_____)
```

::::::::::::::: solution

Step 1 - Install package

```{r}
install.packages('ggplot2')
```

Step 2 - Load package

```{r}
library(ggplot2)
```
:::::::::::::::
::::::::::::::::::::::::::::


